---
title: "exercise-11"
author: "AR"
format: html
editor: visual
---

## Loading in the data set:

```{r}
library(tidyverse)
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f)
```

```{r}
glimpse(d)
skimr::skim(d)
```

*Which of the variables are categorical and which are numeric?*

Categorical variables: order, family, Genus, species

Numeric variables: mass (g), gestation (mo), newborn (g), weaning (mo), wean mass (g), AFR (mo), max. life (mo), litter size, litters/year, refs

## Challenge

*Replace all values of -999 (the authors' code for missing data) with NA.*

```{r}
d[d == -999] <- NA
```

*Drop the variables litter size and refs.*

```{r}
d <- d |>
  select(-refs, -`litter size`)
```

*Rename the variable max.life(mo) to maxlife(mo).*

*Rename the variable wean mass (g) to weanmass(g).*

Note: I actually used the names maxlife_mo and weanmass_g because R was interpreting the parentheses as functions.

Second note! When I tried to run the lms below, R thought the other variables were also functions so I renamed all of them in the code chunk below.

```{r}
library(tidyverse)
d <- d |>
  rename(
    maxlife_mo = `max. life(mo)`, 
    weanmass_g = `wean mass(g)`,
    gestation_mo = `gestation(mo)`,
    weaning_mo = `weaning(mo)`,
    AFR_mo = `AFR(mo)`,
    newborn_g = `newborn(g)`,
    mass_g = `mass(g)`
    )
```

*Log transform all the other numeric variables.*

```{r}
d <- d |>
  mutate(across(where(is.double), log))
```

*Regress the age \[gestation(mo), weaning(mo), AFR(mo), and maxlife(mo)\] and mass \[newborn(g) and weanmass(g)\] variables on overall body mass(g) and add the residuals to the dataframe as new variables \[relGest, relWean, relAFR, relNewbornMass, and relWeaningMass\].*

```{r}
m1 <- lm(data = d, mass_g ~ gestation_mo, na.action = na.exclude)
m2 <- lm(data = d, mass_g ~ weaning_mo, na.action = na.exclude)
m3 <- lm(data = d, mass_g ~ AFR_mo, na.action = na.exclude)
m4 <- lm(data = d, mass_g ~ maxlife_mo, na.action = na.exclude)
m5 <- lm(data = d, mass_g ~ newborn_g, na.action = na.exclude)
m6 <- lm(data = d, mass_g ~ weanmass_g, na.action = na.exclude)

d <- d |>
  mutate(
    relGest = resid(m1),
    relWean = resid(m2),
    relAFR = resid(m3),
    relLife = resid(m4),
    relNewbornMass = resid(m5),
    relWeaningMass = resid(m6)
  )
```

*Plot residuals of max lifespan in relation to Order. Which mammalian orders have the highest residual lifespan?*

The orders with the highest residual lifespans are Sirenia and Perissodactyla. The orders with the highest median residual lifespans are Perissodactyla and Cetacea.

```{r}
p1 <- ggplot(data = d, aes(x = order, y = relLife)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
p1
```

*Plot residuals of newborn mass in relation to Order. Which mammalian orders have the highest residual newborn mass?*

The orders with the highest residuals of newborn mass are Carnivora and Xenarthra. The orders with the highest median residuals of newborn mass are Carnivora and Proboscidea.

```{r}
p2 <- ggplot(data = d, aes(x = order, y = relNewbornMass)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
p2
```

*Plot residuals of weaning mass in relation to Order. Which mammalian orders have the highest residual weaning mass?*

The orders with the highest residuals of weaning mass are Carnivora and Xenarthra. The orders with the highest median residuals of weaning mass are Carnivora and Proboscidea.

```{r}
p3 <- ggplot(data = d, aes(x = order, y = relWeaningMass)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
p3
```

*Run models and a model selection process to evaluate what variables best predict each of the two response variables, maxlife(mo) and AFR(mo), from the set of the following predictors: gestation(mo), newborn(g), weaning(mo), weanmass(g), litters/year, and overall body mass(g).*

Winnowing data set:

```{r}
d <- d |>
  drop_na(maxlife_mo) |>
  drop_na(AFR_mo) |>
  drop_na(gestation_mo) |>
  drop_na(newborn_g) |>
  drop_na(weaning_mo) |>
  drop_na(weanmass_g) |>
  drop_na(`litters/year`) |>
  drop_na(mass_g)
```

Run models and model selection process:

**Maxlife(mo):**

```{r}
library(MuMIn)
m_full_ml <- lm(data = d, maxlife_mo ~ gestation_mo + newborn_g + weaning_mo + weanmass_g + `litters/year` + mass_g, na.action = na.fail)
mods_ml <- dredge(m_full_ml)
mods_ml
```

```{r}
mods_ml_avg <- summary(model.avg(mods_ml, subset = delta <= 4, fit = TRUE))
mods_ml_avg$msTable
confint(mods_ml_avg)
plot(mods_ml_avg, full = TRUE)
```

**AFR(mo):**

```{r}
m_full_afr <- lm(data = d, AFR_mo ~ gestation_mo + newborn_g + weaning_mo + weanmass_g + `litters/year` + mass_g, na.action = na.fail)
mods_afr <- dredge(m_full_afr)
mods_afr
```

```{r}
mods_afr_avg <- summary(model.avg(mods_afr, subset = delta <= 4, fit = TRUE))
mods_afr_avg$msTable
confint(mods_afr_avg)
plot(mods_afr_avg, full = TRUE)
```

*For each of the two response variables, indicate what is the best model overall based on AICc and how many models have a delta AICc of 4 or less?*

**Maxlife(mo):** The best model overall uses the variables litters/year, gestation(mo), mass(g), and weaning(mo) to predict maxlife(mo). It has an AICc of 256.83. There are four other models that have a delta AICc of 4 or less, with AICc values 257.86, 258.95, 259.67, and 259.91.

**AFR(mo):** The best model overall uses the variables litters/year, gestation(mo), mass(g), and weaning(mo) to predict age at first repoduction(mo). It has an AICc of 321.23. There are nine other models that have a delta AICc of 4 or less, with AICc values 321.52, 321.82, 321,82, 323.37, 323.37, 323.65, 324.12, 324.81, and 324.94.

*What variables, if any, appear in all of this set of "top" models?*

**Maxlife(mo):** The variables that appear in all of this set of top models are 1, 2, and 5, litters/year, gestation(mo) and weaning(mo), respectively.

**AFR(mo):** The variables that appear in all of this set of top models are variables 1 and 2, litters/year and gestation(mo), respectively.

*Calculate and plot the model-averaged coefficients and their CIs across this set of top models.*

Plotted above.
